---
title: "Machine Learning Project"
author: "jp"
date: "October 20, 2014"
output: html_document
---

Machine Learning Project
Step-by-step explanation and documentation, from training data to test data.

STEPS:
Create folder if needed
```{r}
if (!file.exists("data")) {dir.create("data")}
```
URL and placement
```{r}
fileUrl1 <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
destfile1 <- "./data/pml-training.csv"
fileUrl2 <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
destfile2 <- "./data/pml-testing.csv"
```
Download and Date
```{r}
download.file(fileUrl1, destfile = destfile1)
download.file(fileUrl2, destfile = destfile2)
dateDownloaded <- date()
```
Read csv Training Data
```{r}
data_training <- read.csv("./data/pml-training.csv", na.strings= c("NA",""," "))
```
Remove NAs
```{r}
data_training_NAs <- apply(data_training, 2, function(x) {sum(is.na(x))})
data_training_clean <- data_training[,which(data_training_NAs == 0)]
```
Remove Unnecessary Columns
```{r}
data_training_clean <- data_training_clean[8:length(data_training_clean)]
```
Required Packages
```{r}
require(caret)
require(corrplot)
require(randomForest)
require(kernlab)
```
Split for Cross-Validation
```{r}
inTrain <- createDataPartition(y = data_training_clean$classe, p = 0.7, list = FALSE)
training <- data_training_clean[inTrain, ]
crossval <- data_training_clean[-inTrain, ]
```
Plot Correlation Matrix
```{r, echo = FALSE}
correlMatrix <- cor(training[, -length(training)])
corrplot(correlMatrix, order = "FPC", method = "circle", type = "lower", tl.cex = 0.8,  tl.col = rgb(0, 0, 0))
```
Model classe ~ ALL
```{r}
(model <- randomForest(classe ~ ., data = training))
```
Cross-Validate
```{r}
predictCrossVal <- predict(model, crossval)
(confusionMatrix(crossval$classe, predictCrossVal))
```
Redo with Testing Data
```{r}
data_test <- read.csv("./data/pml-testing.csv", na.strings= c("NA",""," "))
data_test_NAs <- apply(data_test, 2, function(x) {sum(is.na(x))})
data_test_clean <- data_test[,which(data_test_NAs == 0)]
data_test_clean <- data_test_clean[8:length(data_test_clean)]
```
Predict classe | Test Data
```{r}
(predictTest <- predict(model, data_test_clean))
```